
By : Khan Akbar 


Title: Project 1: Exploring Weather Trends  
Author: khan akbar 
Date: 01-05-2020 
Output: project-1.pdf  


## Outline
* Extract the data
* Open up the CSV
* Create a line chart
* Make four observations


### Extract the data
- city list  
    SELECT *  
    FROM city_list  
    ORDER BY country, city  

- city data  
    SELECT *  
    FROM city_data  
    WHERE city IN ('San Jose', 'Los Angeles','San Francisco') AND country = 'United States'  

- global data  
    SELECT *  
    FROM global_data    



Data import the data in R Studio 
```{r message=TRUE, warning=TRUE, paged.print=TRUE}
## 1) Reading datasets
library(readr)
library(dplyr)

# a. Global 
global <- read_csv("C:/Users/khana/OneDrive/Desktop/UDACITY project/Udacity-DAND-Weather-Trends-master (1)/Udacity-DAND-Weather-Trends-master/data/global.csv")

# Make the time period the same like San Jose and Los Angeles
global2 <- global %>%
  filter(year >= 1849 & year <= 2013)

# b. My current Location: San Jose
cities <- read_csv("C:/Users/khana/OneDrive/Desktop/UDACITY project/Udacity-DAND-Weather-Trends-master (1)/Udacity-DAND-Weather-Trends-master/data/cities.csv")
sj <- cities %>%
  filter(city == 'San Jose') %>%
  select(year,avg_temp)

# c. The city I stayed before the Bay: Los Angeles
la <- cities %>%
  filter(city == 'Los Angeles') %>%
  select(year,avg_temp)

# d. My hometown: Taipei
tpi <- read_csv("C:/Users/khana/OneDrive/Desktop/UDACITY project/Udacity-DAND-Weather-Trends-master (1)/Udacity-DAND-Weather-Trends-master/data/tpi.csv") %>%
  filter(year >= 1849 & year <= 2013) %>%
  select(year,avg_temp)

```



Calculating moving average
```{r}
##################################### 2) Create a function for calculating moving average ###############################3
move <- function(data,loc,y){
    mov_year = data.frame()
    locat = data.frame()
    year = data.frame()
    movavg = data.frame()
    for(i in 1:length(data$avg_temp)){
        if (i < y){
            x <- NA
            z <- NA
        }
            else{
                x <- data %>%
                summarize(movmean = mean(avg_temp[(i-(y-1)):i]))
                z <- paste(y,'year_moving_avg')
                }
        movavg = rbind(movavg,data.frame(as.numeric(x)))
        year = rbind(year,z)
        locat = rbind(locat,loc)
    }
    names(locat) <- 'location'
    names(movavg) <- 'mov_avg'
    names(year) = 'temp_moving_avg'
    mov_year <- cbind(data,movavg,year,locat)
    return(mov_year)
}

globalall <- rbind(move(global2,'Global',3),move(global2,'Global',5),move(global2,'Global',7),move(global2,'Global',10))

library(ggplot2)
ggplot(data=globalall, aes(x=year, y=mov_avg, color=temp_moving_avg))+geom_line()+
  ylab("Average Temperature")+theme_bw()+ggtitle("Compare among different temperature moving average")
```


Removing 1st to 6th year (NAs) 
```{r}
# Remove the missing value and data prepreation by remob=ving  NAs
global7 <- move(global2,'Global',7) %>%
    filter(year >= 1855)
head(global7)

sj7 <- move(sj,'San Jose',7) %>%
  filter(year >= 1855)
head(sj7)

glo_sj <- rbind(global7,sj7)

la7 <- move(la,'Los Angeles',7) %>%
  filter(year >= 1855)
head(la7)

tpi7 <- move(tpi,'Taipei',7) %>% 
  filter(year >= 1855)
head(tpi7)

glob_city <- rbind(global7,sj7,la7,tpi7)
```


Compare weather trends in global and cities 
```{r}
## 3) Data Visualization
library(ggplot2)

ggplot(data=global7, aes(x=year, y=mov_avg, color=temp_moving_avg))+geom_line()+
    ylab("Average Temperature")+theme_bw()+ggtitle("7 year Moving Average Temperature in Global")

ggplot(data=sj7, aes(x=year, y=mov_avg, color=temp_moving_avg))+geom_line()+
  ylab("Average Temperature")+theme_bw()+ggtitle("7 year Moving Average Temperature in San Jose")

ggplot(data=glo_sj, aes(x=year, y=mov_avg, color=location))+geom_line()+
  ylab("Average Temperature")+theme_bw()+ggtitle("7 year Moving Average Temperature Global V.S. San Jose")

ggplot(data=la7, aes(x=year, y=mov_avg, color=temp_moving_avg))+geom_line()+
  ylab("Average Temperature")+theme_bw()+ggtitle("7 year Moving Average Temperature in Los Angeles")

ggplot(data=tpi7, aes(x=year, y=mov_avg, color=temp_moving_avg))+geom_line()+
  ylab("Average Temperature")+theme_bw()+ggtitle("7 year Moving Average Temperature in Taipei")

ggplot(data=glob_city, aes(x=year, y=mov_avg, color=location))+geom_line()+
  ylab("Average Temperature")+theme_bw()+ggtitle("7 year Moving Average Temperature")
```

